---
title: "Data Analysis Project Proposal"
date: "07/19/2020"
author: "STAT 420, Summer 2020"
output:
  html_document: 
    theme: readable
    toc: yes
urlcolor: cyan
---

*** 

```{r, message=FALSE, warning=FALSE}
library(readr)
dat = read_csv("new.csv")
```

*** 

# Team Members

- Chenyue Jiao (cjiao4)
- Xiaojuan Wang
- Yan Sun (yansun5)
- Yulin Li (yulinl2)

*** 

# Tentative Project Title

**Prediction of Housing Price in Beijing by Multiple Linear Regression Modeling**

*** 

# Data Description

We are going to use the dataset about housing prices in Beijing from 2011 to 2017. It has a total of `r nrow(dat)` observations. If we remove those observations with `NaN` values, we still have `r nrow(subset(dat, buildingType != "NaN"))` observations left. This dataset has `r ncol(dat)` variables. Among them, we will use `price` as the response variable, which represents the average price (in Chinese RMB) per square meter. As for the predictor variables, we plan to use the followings:

- `livingRoom`: the number of living room. numeric predictor.
- `drawingRoom`: the number of drawing room. numeric predictor.
- `kitchen`: the number of kitchen. numeric predictor.
- `bathroom`: the number of bathroom. numeric predictor.
- `floor`: the floor that the house is on. numeric predictor. (this predictor has some messy code in it, which we will preprocess before using it.)
- `buildingType`: the type of building. categorical predictor, including tower (1), bungalow (2), combination of plate and tower (3), plate (4).
- `constructionTime`: the time of construction. numeric predictor.
- `renovationCondition`: the condition of renovation. categorical predictor, including other (1), rough (2), simplicity (3), hardcover (4).
- `buildingStructure`: the structure type of the building. categorical predictor, including unknown (1), mixed (2), brick and wood (3), brick and concrete (4), steel (5), steel-concrete composite (6).
- `ladderRatio`: the proportion between number of residents on the same floor and number of elevator of ladder. It describes how many ladders a resident have on average. numeric predictor.
- `elevator`: whether the building has elevators (1) or not (0). categorical predictor.
- `subway`: whether there is (1) subways nearby or not (0). categorical predictor.
- `district`: which district the house is located in Beijing. categorical predictor including 13 levels.

*** 

# Backgroud Information

We obtained our dataset from [this page](https://www.kaggle.com/ruiqurm/lianjia) on Kaggle. All the data in this dataset was fetched by its author Qichen Liu from [this page](https://bj.lianjia.com/chengjiao) on the Chinese housing website [Lianjia.com](https://bj.lianjia.com).

*** 

# Statement of Interest

Over the past decade, rising prices in China's housing market have been increasingly overwhelming to the young families trying to settle down in major cities in China. As a group of university students potentially facing this issue in our near future, we want to understand how much housing prices in a Chinese metropolis like Beijing can be predicted and explained by some visible factors in the market, for example, characteristics of a house itself. 

In view of what we have learned in this course, we set our goal as building a multiple regression model capable of making predictions of the housing price with hope that this model could also provide us with further insights into the housing market. 

*** 

# Snapshot of the Response Variable

```{r}
head(dat$price)
```

*** 



