---
title: "Covid-19 Policy Helper"
date: Aug 2020
author: "Xinyi Lai, Yulin Li"
output: 
  html_document: 
    theme: readable
    toc: yes
    toc_depth: 4
    number_sections: no
urlcolor: cyan
editor_options: 
  chunk_output_type: inline
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80)
library(knitr)
opts_chunk$set(cache = TRUE, autodep = TRUE)
```

## Introduction

This is `part 3` of the project "Covid-19 Policy Decision Helper". In this part, we try to model the relationship between the SEIR model parameters and a variety of social factors, including demographic, medical and policy factors. We aim at obtaining models that are interpretive as well as predictive; in order words, we are hoping to find models that are simple, accessible and easy to be interpreted and understood, so that people can gain some insights of what is significant to the way a pandemic develops, but at the same time, we are also striving to find models that are most helpful making predictions among those explainable models. 

With such goals in mind, we engaged a relatively small number of variables in our study--variables that seem significant to us intuitively, from the most accessible open data source. 

The **modeling methods** we applied include the following: 

- Data cleaning as necessary to address observations with missing or extreme values.
- Multiple linear regression
- ANOVA (codes omitted)
- Interaction (codes omitted)
- Residual diagnostics
- Transformations
- Polynomial regression
- Stepwise model selection (AIC & BIC) (codes omitted)
- Variable selection (codes omitted)
- Test/train splitting

**Data source**: 

- Demographic, medical and policy data: [COVID-19 Prediction Models Counties & Hospitals | Yu Group (UC Berkeley)](https://aws.amazon.com/marketplace/pp/prodview-px2tvvydirx4o?qid=1587582026402&sr=0-1&ref_=srh_res_product_title#overview)
- Infection data and SEIR model estimated parameters: project `part 1` and `part 2`.

```{r echo=FALSE, message=FALSE}
library(lmtest)

get_bp_decision = function(model, alpha) {
  decide = unname(bptest(model)$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_bp_pval = function(model) {
  bptest(model)$p.value
}

get_sw_decision = function(model, alpha) {
  decide = unname(shapiro.test(resid(model))$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_sw_pval = function(model) {
  shapiro.test(resid(model))$p.value
}

get_num_params = function(model) {
  length(coef(model))
}

get_loocv_rmse = function(model, is_log, k) {
  ifelse(
    is_log, 
    sqrt(mean(na.omit(((data_trn[, k] - exp(fitted(model))) / (1 - hatvalues(model))) ^ 2))),
    sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
  )
}

get_adj_r2 = function(model) {
  summary(model)$adj.r.squared
}

test_mod = function(model, is_log = FALSE, k = 1){
  c(loocv_rmse = get_loocv_rmse(model, is_log, k), 
    adj_r2 = get_adj_r2(model), 
    bp_pval = get_bp_pval(model), 
    sw_pval = get_sw_pval(model), 
    num_params = get_num_params(model), 
    test_rmse = get_test_rmse(model, k), 
    perc_err = get_perc_err(model, k))
}

diagnostics = function(model, pcol = "grey", lcol = "dodgerblue", alpha = 0.05, plotit = TRUE, testit = TRUE){
  if (plotit){
    par(mfrow = c(1, 2), pty="s")
    
    plot(fitted(model), resid(model), col = "grey", pch = 20, 
         xlab = "Fitted", ylab = "Residual", 
         main = "Fitted versus Residuals")
    abline(h = 0, col = "darkorange", lwd = 2)
    
    qqnorm(resid(model), col = pcol)
    qqline(resid(model), col = lcol, lwd = 2)
  }
  if (testit){
    list(p_val = shapiro.test(resid(model))$p, 
         decision = ifelse(test = shapiro.test(resid(model))$p < alpha, 
                           yes = "Reject", no = "Fail to Reject"))
  }
}

get_test_rmse = function(model, k) {
  sqrt(mean((data_tst[, k] - predict(model, newdata = data_tst))^ 2))
}

get_perc_err = function(model, k) {
  actual = data_tst[, k]
  predicted = predict(model, newdata = data_tst)
  100 * mean((abs(actual - predicted)) / actual)
}
```

## Data Preprocessing

```{r}
StateOfInterest = c("Arizona", "California", "Minnesota", "New Mexico", "New York", 
                   "Oklahoma", "South Carolina", "Tennessee", "Utah", "Virginia", 
                   "West Virginia", "Wisconsin")
```


### Demographic & Policy Data (County-Level)
```{r message=FALSE}
library(readr)
county_data_abridged = read_csv("../datasets/county_data_abridged.csv")
```

```{r}
dim(county_data_abridged)
names(county_data_abridged)
```



```{r}
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "#EligibleforMedicare2018"] = "EligibleforMedicare2018"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "#FTEHospitalTotal2017"] = "FTEHospitalTotal2017"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "#HospParticipatinginNetwork2017"] = "HospParticipatinginNetwork2017"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "#Hospitals"] = "Hospitals"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "#ICU_beds"] = "ICU_beds"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "PopulationEstimate65+2017"] = "PopulationEstimate_above65_2017"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "stay at home"] = "stay_at_home"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == ">50 gatherings"] = "above_50_gatherings"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == ">500 gatherings"] = "above_500_gatherings"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "public schools"] = "public_schools"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "restaurant dine-in"] = "restaurant_dine_in"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "entertainment/gym"] = "entertainment_gym"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "federal guidelines"] = "federal_guidelines"
colnames(county_data_abridged)[colnames(county_data_abridged) 
                               == "foreign travel ban"] = "foreign_travel_ban"
```


```{r}
data = subset(county_data_abridged,
              select = c(State, CountyName, POP_LATITUDE, POP_LONGITUDE,
                         PopulationEstimate2018, PopTotalMale2017, 
                         PopulationEstimate_above65_2017, PopulationDensityperSqMile2010, 
                         DiabetesPercentage, Smokers_Percentage, 
                         HeartDiseaseMortality, StrokeMortality, 
                         Hospitals, ICU_beds, HospParticipatinginNetwork2017, 
                         stay_at_home, above_50_gatherings, above_500_gatherings, 
                         restaurant_dine_in, entertainment_gym))
data = na.omit(data)
data = droplevels(data)

data$stay_at_home = data$stay_at_home - range(data$stay_at_home)[1]
data$above_50_gatherings = data$above_50_gatherings - range(data$above_50_gatherings)[1]
data$above_500_gatherings = data$above_500_gatherings - range(data$above_500_gatherings)[1]
data$restaurant_dine_in = data$restaurant_dine_in - range(data$restaurant_dine_in)[1]
data$entertainment_gym = data$entertainment_gym - range(data$entertainment_gym)[1]

str(data)
```

```{r}
data_demographic_county = data
```


### Infection Data (State-Level)

#### Covid-19 Cases
```{r message=FALSE, warning=FALSE}
timeseries = read_csv("../datasets/timeseries.csv")
```

```{r}
data = timeseries
data = subset(data, country == "United States" & level == "state")
data = subset(data, !(name %in% c("Unassigned cases, Arkansas, US", 
              "Unassigned cases, Georgia, US", "Unassigned cases, Illinois, US",
              "Unassigned cases, Iowa, US", "Unassigned cases, Maine, US",
              "Unassigned cases, Massachusetts, US", "Unassigned cases, North Dakota, US",
              "Washington, D.C., US")))
data$state = matrix(unlist(strsplit(as.character(data$name), ", ")), ncol = 2, byrow = TRUE)[, 1]
data = subset(data,
              select = c(state, date, cases, deaths, recovered))

data = subset(data, state %in% StateOfInterest)

data$state =     as.factor(data$state)
data$cases =     as.numeric(data$cases)
data$deaths =    as.numeric(data$deaths)
data$recovered = as.numeric(data$recovered)

data = na.omit(data)
data = droplevels(data)

str(data)
```

```{r}
range(data$cases)
range(data$deaths)
range(data$recovered)
```

```{r}
data_cases_state = data
```

#### SEIR Model Parameters
```{r message=FALSE, warning=FALSE}
model_out = read_csv("../datasets/model_out.csv")
```

```{r}
str(model_out)
```

```{r}
data_parm = model_out
```

### Combined Data of Intersts (State-Level)

```{r}
data = data_parm

n = length(data$state)
data$cases = rep(0, n)
data$deaths = rep(0, n)
data$recovered = rep(0, n)

for (i in 1:n){
  j = data_cases_state$state == data$state[i] & data_cases_state$date == data$startdate[i]
  if (sum(j) == 0){
    data[i, ] = NA
  }else{
    data[i, c("cases", "deaths", "recovered")] = data_cases_state[j, c("cases", "deaths", "recovered")]
  }
}

data = na.omit(data)

data_demographic_state = as.data.frame(matrix(nrow = length(data$state), ncol = length(colnames(data_demographic_county)) - 1))
colnames(data_demographic_state) = colnames(data_demographic_county)[colnames(data_demographic_county) != "CountyName"]
data_demographic_state$State = data$state

for (s in data_demographic_state$State){
  for (k in 2:ncol(data_demographic_state)){
    data_demographic_state[data_demographic_state$State == s, k] = 
      mean(unlist(data_demographic_county[data_demographic_county$State == s, k+1]))
  }
}

data = cbind(data, 
             subset(data_demographic_state[, colnames(data_demographic_state)[colnames(data_demographic_state) != "State"]]))
data = na.omit(data)
```

```{r}
data$days = rep(0, nrow(data))
data$temp = rep(0, nrow(data))
for (i in 1:nrow(data)){
  data$temp = diff.Date(c(data[i, ]$startdate, data[i, ]$enddate))
}
for (s in data$state){
  for (d in 1:nrow(subset(data, state == s))){
    data[data$state == s, "days"][d] = sum(subset(data, state == s)[1:d, "temp"])
  }
}
data = data[, colnames(data)[colnames(data) != "temp"]]
data_unamed = data[, colnames(data)[!(colnames(data) %in% c("state", "startdate", "enddate"))]]
```


```{r}
# head(data_demographic_county, 10)
# head(data_demographic_state, 10)
# head(data[, c("state", "startdate", "enddate", "days")], 10)
head(data, 10)
```



## Methods: Modeling & Testing Procedures (abridged)

### Data Points `pairs()` Plot

<style>
  .superbigimage{
      overflow-x:scroll;
      white-space: nowrap;
  }

  .superbigimage img{
     max-width: none;
  }
  
  
</style>

<div class="superbigimage">
```{r, echo=FALSE, fig.width=30, fig.height=20}
pairs(data_unamed, col = "dodgerblue")
```
</div>

```{r eval=FALSE, include=FALSE}
colnames(data_unamed)
```

```{r}
set.seed(42)
num_obs = nrow(data_unamed) # total number of observations
num_trn = round(num_obs * 0.90) # number of observations for the training data

trn_idx = sample(num_obs, num_trn) # randomly generate the index for the training data
data_trn = data_unamed[trn_idx, ] # training data
data_tst = data_unamed[-trn_idx, ] # testing data
```

### Modeling `k`
```{r}
# full additive model
mod_k_full = lm(k ~ ., data = data_trn)
test_mod(mod_k_full, k = 1)
summary(mod_k_full)
```


```{r}
# small additive model
mod_k_1 = lm(k ~ days, data = data_trn)
# large additive model
mod_k_2 = lm(k ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma, data = data_trn)

test_mod(mod_k_1, k = 1)
test_mod(mod_k_2, k = 1)
summary(mod_k_1)
summary(mod_k_2)
```
```{r}
# intermediate model
mod_k_3 = lm(k ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 
             + log(days) - days
             + stay_at_home - PopulationEstimate2018 - PopTotalMale2017
             + I(POP_LATITUDE ^ 2) +  + I(POP_LONGITUDE ^ 2), 
             data = data_trn)
mod_k_4 = lm(k ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 + 
               log(days) - days, 
             data = data_trn)

test_mod(mod_k_1, k = 1)
test_mod(mod_k_2, k = 1)
test_mod(mod_k_3, k = 1)
test_mod(mod_k_4, k = 1)
summary(mod_k_3)
summary(mod_k_4)
```

```{r}
# intermediate model 
# relatively bad models
mod_k_5 = lm(k ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - (lamda + c + alpha + omega + miu), 
             data = data_trn)
mod_k_6 = lm(k ~ lamda + c + alpha + omega + miu, 
             data = data_trn)

test_mod(mod_k_1, k = 1)
test_mod(mod_k_2, k = 1)
test_mod(mod_k_3, k = 1)
test_mod(mod_k_4, k = 1)
test_mod(mod_k_5, k = 1)
test_mod(mod_k_6, k = 1)
summary(mod_k_5)
summary(mod_k_6)
```

```{r}
mod_k = mod_k_6
# diagnostics(mod_k, testit = FALSE)
```

### Modeling `sigma`

In this case, `sigma` is set to an empirical value. Modeling excluded.

### Modeling `lamda`

```{r}
# full additive model
mod_lamda_full = lm(lamda ~ ., data = data_trn)
test_mod(mod_lamda_full, k = 3)
summary(mod_lamda_full)
```


```{r}
# small additive model
mod_lamda_1 = lm(lamda ~ days, data = data_trn)
# large additive model
mod_lamda_2 = lm(lamda ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma, data = data_trn)

test_mod(mod_lamda_1, k = 3)
test_mod(mod_lamda_2, k = 3)
summary(mod_lamda_1)
summary(mod_lamda_2)
```

```{r}
# intermediate model
mod_lamda_3 = lm(lamda ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 
             + log(days) - days
             + stay_at_home - PopulationEstimate2018 - PopTotalMale2017
             + I(POP_LATITUDE ^ 2) +  + I(POP_LONGITUDE ^ 2), 
             data = data_trn)
mod_lamda_4 = lm(lamda ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 + 
               log(days) - days, 
             data = data_trn)

test_mod(mod_lamda_1, k = 3)
test_mod(mod_lamda_2, k = 3)
test_mod(mod_lamda_3, k = 3)
test_mod(mod_lamda_4, k = 3)
summary(mod_lamda_3)
summary(mod_lamda_4)
```

```{r}
# intermediate model 
# relatively bad models
mod_lamda_5 = lm(lamda ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - (k + c + alpha + omega + miu), 
             data = data_trn)
mod_lamda_6 = lm(lamda ~  k + c + alpha + omega + miu, 
             data = data_trn)

test_mod(mod_lamda_1, k = 3)
test_mod(mod_lamda_2, k = 3)
test_mod(mod_lamda_3, k = 3)
test_mod(mod_lamda_4, k = 3)
test_mod(mod_lamda_5, k = 3)
test_mod(mod_lamda_6, k = 3)
summary(mod_lamda_5)
summary(mod_lamda_6)
```

```{r}
mod_lamda = mod_lamda_4
# diagnostics(mod_lamda, testit = FALSE)
```


### Modeling `c`

```{r}
# full additive model
mod_c_full = lm(c ~ ., data = data_trn)
test_mod(mod_c_full, k = 4)
summary(mod_c_full)
```


```{r}
# small additive model
mod_c_1 = lm(c ~ days, data = data_trn)
# large additive model
mod_c_2 = lm(c ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma, data = data_trn)

test_mod(mod_c_1, k = 4)
test_mod(mod_c_2, k = 4)
summary(mod_c_1)
summary(mod_c_2)
```

```{r}
# intermediate model
mod_c_3 = lm(c ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 
             + log(days) - days
             + stay_at_home - PopulationEstimate2018 - PopTotalMale2017
             + I(POP_LATITUDE ^ 2) +  + I(POP_LONGITUDE ^ 2), 
             data = data_trn)
mod_c_4 = lm(c ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 + 
               log(days) - days, 
             data = data_trn)

test_mod(mod_c_1, k = 4)
test_mod(mod_c_2, k = 4)
test_mod(mod_c_3, k = 4)
test_mod(mod_c_4, k = 4)
summary(mod_c_3)
summary(mod_c_4)
```

```{r}
# intermediate model 
# relatively bad models
mod_c_5 = lm(c ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - (k + lamda + alpha + omega + miu), 
             data = data_trn)
mod_c_6 = lm(c ~  k + lamda + alpha + omega + miu, 
             data = data_trn)

test_mod(mod_c_1, k = 4)
test_mod(mod_c_2, k = 4)
test_mod(mod_c_3, k = 4)
test_mod(mod_c_4, k = 4)
test_mod(mod_c_5, k = 4)
test_mod(mod_c_6, k = 4)
summary(mod_c_5)
summary(mod_c_6)
```

```{r}
mod_c = mod_c_3
# diagnostics(mod_c, testit = FALSE)
```

### Modeling `alpha`

```{r}
# full additive model
mod_alpha_full = lm(alpha ~ ., data = data_trn)
test_mod(mod_alpha_full, k = 5)
summary(mod_alpha_full)
```


```{r}
# small additive model
mod_alpha_1 = lm(alpha ~ days, data = data_trn)
# large additive model
mod_alpha_2 = lm(alpha ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma, data = data_trn)

test_mod(mod_alpha_1, k = 5)
test_mod(mod_alpha_2, k = 5)
summary(mod_alpha_1)
summary(mod_alpha_2)
```

```{r}
# intermediate model
mod_alpha_3 = lm(alpha ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 
             + log(days) - days
             + stay_at_home - PopulationEstimate2018 - PopTotalMale2017
             + I(POP_LATITUDE ^ 2) +  + I(POP_LONGITUDE ^ 2), 
             data = data_trn)
mod_alpha_4 = lm(alpha ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 + 
               log(days) - days, 
             data = data_trn)

test_mod(mod_alpha_1, k = 5)
test_mod(mod_alpha_2, k = 5)
test_mod(mod_alpha_3, k = 5)
test_mod(mod_alpha_4, k = 5)
summary(mod_alpha_3)
summary(mod_alpha_4)
```

```{r}
# intermediate model 
# relatively bad models
mod_alpha_5 = lm(alpha ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - (k + lamda + c + omega + miu), 
             data = data_trn)
mod_alpha_6 = lm(alpha ~  k + lamda + c + omega + miu, 
             data = data_trn)

test_mod(mod_alpha_1, k = 5)
test_mod(mod_alpha_2, k = 5)
test_mod(mod_alpha_3, k = 5)
test_mod(mod_alpha_4, k = 5)
test_mod(mod_alpha_5, k = 5)
test_mod(mod_alpha_6, k = 5)
summary(mod_alpha_5)
summary(mod_alpha_6)
```

```{r}
mod_alpha = mod_alpha_4
# diagnostics(mod_alpha, testit = FALSE)
```

### Modeling `omega`

```{r}
# full additive model
mod_omega_full = lm(omega ~ ., data = data_trn)
test_mod(mod_omega_full, k = 6)
summary(mod_omega_full)
```


```{r}
# small additive model
mod_omega_1 = lm(omega ~ days, data = data_trn)
# large additive model
mod_omega_2 = lm(omega ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma, data = data_trn)

test_mod(mod_omega_1, k = 6)
test_mod(mod_omega_2, k = 6)
summary(mod_omega_1)
summary(mod_omega_2)
```

```{r}
# intermediate model
mod_omega_3 = lm(omega ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 
             + log(days) - days
             + stay_at_home - PopulationEstimate2018 - PopTotalMale2017
             + I(POP_LATITUDE ^ 2) +  + I(POP_LONGITUDE ^ 2), 
             data = data_trn)
mod_omega_4 = lm(omega ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 + 
               log(days) - days, 
             data = data_trn)

test_mod(mod_omega_1, k = 6)
test_mod(mod_omega_2, k = 6)
test_mod(mod_omega_3, k = 6)
test_mod(mod_omega_4, k = 6)
summary(mod_omega_3)
summary(mod_omega_4)
```

```{r}
# intermediate model 
# relatively bad models
mod_omega_5 = lm(omega ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - (k + lamda + c + alpha + miu), 
             data = data_trn)
mod_omega_6 = lm(omega ~  k + lamda + c + alpha + miu, 
             data = data_trn)

test_mod(mod_omega_1, k = 6)
test_mod(mod_omega_2, k = 6)
test_mod(mod_omega_3, k = 6)
test_mod(mod_omega_4, k = 6)
test_mod(mod_omega_5, k = 6)
test_mod(mod_omega_6, k = 6)
summary(mod_omega_5)
summary(mod_omega_6)
```

```{r}
mod_omega = mod_omega_4
# diagnostics(mod_omega, testit = FALSE)
```

### Modeling `miu`

```{r}
# full additive model
mod_miu_full = lm(miu ~ ., data = data_trn)
test_mod(mod_miu_full, k = 7)
summary(mod_miu_full)
```


```{r}
# small additive model
mod_miu_1 = lm(miu ~ days, data = data_trn)
# large additive model
mod_miu_2 = lm(miu ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma, data = data_trn)

test_mod(mod_miu_1, k = 7)
test_mod(mod_miu_2, k = 7)
summary(mod_miu_1)
summary(mod_miu_2)
```

```{r}
# intermediate model
mod_miu_3 = lm(miu ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 
             + log(days) - days
             + stay_at_home - PopulationEstimate2018 - PopTotalMale2017
             + I(POP_LATITUDE ^ 2) +  + I(POP_LONGITUDE ^ 2), 
             data = data_trn)
mod_miu_4 = lm(miu ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - cases - deaths - omega - recovered - PopulationDensityperSqMile2010 + 
               log(days) - days, 
             data = data_trn)

test_mod(mod_miu_1, k = 7)
test_mod(mod_miu_2, k = 7)
test_mod(mod_miu_3, k = 7)
test_mod(mod_miu_4, k = 7)
summary(mod_miu_3)
summary(mod_miu_4)
```

```{r}
# intermediate model 
# relatively bad models
mod_miu_5 = lm(miu ~ . - Hospitals - ICU_beds - HospParticipatinginNetwork2017 - stay_at_home - above_50_gatherings - above_500_gatherings - restaurant_dine_in - entertainment_gym - sigma
             - (k + lamda + c + alpha + omega), 
             data = data_trn)
mod_miu_6 = lm(miu ~  k + lamda + c + alpha + omega, 
             data = data_trn)

test_mod(mod_miu_1, k = 7)
test_mod(mod_miu_2, k = 7)
test_mod(mod_miu_3, k = 7)
test_mod(mod_miu_4, k = 7)
test_mod(mod_miu_5, k = 7)
test_mod(mod_miu_6, k = 7)
summary(mod_miu_5)
summary(mod_miu_6)
```

```{r}
mod_miu = mod_miu_2
# diagnostics(mod_miu, testit = FALSE)
```

## Result

### Six Models & Summaries

```{r}
summary(mod_k)
summary(mod_lamda)
summary(mod_c)
summary(mod_alpha)
summary(mod_omega)
summary(mod_miu)
```

### Model Diagonostics (Test Statistics)

```{r}
test_mod(mod_k, k = 1)
test_mod(mod_lamda, k = 3)
test_mod(mod_c, k = 4)
test_mod(mod_alpha, k = 5)
test_mod(mod_omega, k = 6)
test_mod(mod_miu, k = 7)
```

### Model Diagonostics (Plots)

```{r fig.cap="[Figure 1.1. Model `k`]", fig.align="center"}
diagnostics(mod_k, testit = FALSE)
```


```{r fig.cap="[Figure 1.2. Model `lamda`]", fig.align="center"}
diagnostics(mod_lamda, testit = FALSE)
```


```{r fig.cap="[Figure 1.3. Model `c`]", fig.align="center"}
diagnostics(mod_c, testit = FALSE)
```


```{r fig.cap="[Figure 1.4. Model `alpha`]", fig.align="center"}
diagnostics(mod_alpha, testit = FALSE)
```


```{r fig.cap="[Figure 1.5. Model `omega`]", fig.align="center"}
diagnostics(mod_omega, testit = FALSE)
```


```{r fig.cap="[Figure 1.6. Model `miu`]", fig.align="center"}
diagnostics(mod_miu, testit = FALSE)
```


## Discussion

See project [Homepage](https://xinyi-lai.github.io). 

## Appendix 

### Helper Functions 

```{r helper_functions, eval=FALSE}
get_bp_decision = function(model, alpha) {
  decide = unname(bptest(model)$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_bp_pval = function(model) {
  bptest(model)$p.value
}

get_sw_decision = function(model, alpha) {
  decide = unname(shapiro.test(resid(model))$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_sw_pval = function(model) {
  shapiro.test(resid(model))$p.value
}

get_num_params = function(model) {
  length(coef(model))
}

get_loocv_rmse = function(model, is_log, k) {
  ifelse(
    is_log, 
    sqrt(mean(na.omit(((data_trn[, k] - exp(fitted(model))) / (1 - hatvalues(model))) ^ 2))),
    sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
  )
}

get_adj_r2 = function(model) {
  summary(model)$adj.r.squared
}

test_mod = function(model, is_log = FALSE, k = 1){
  c(loocv_rmse = get_loocv_rmse(model, is_log, k), 
    adj_r2 = get_adj_r2(model), 
    bp_pval = get_bp_pval(model), 
    sw_pval = get_sw_pval(model), 
    num_params = get_num_params(model), 
    test_rmse = get_test_rmse(model, k), 
    perc_err = get_perc_err(model, k))
}

diagnostics = function(model, pcol = "grey", lcol = "dodgerblue", alpha = 0.05, plotit = TRUE, testit = TRUE){
  if (plotit){
    par(mfrow = c(1, 2), pty="s")
    
    plot(fitted(model), resid(model), col = "grey", pch = 20, 
         xlab = "Fitted", ylab = "Residual", 
         main = "Fitted versus Residuals")
    abline(h = 0, col = "darkorange", lwd = 2)
    
    qqnorm(resid(model), col = pcol)
    qqline(resid(model), col = lcol, lwd = 2)
  }
  if (testit){
    list(p_val = shapiro.test(resid(model))$p, 
         decision = ifelse(test = shapiro.test(resid(model))$p < alpha, 
                           yes = "Reject", no = "Fail to Reject"))
  }
}

get_test_rmse = function(model, k) {
  sqrt(mean((data_tst[, k] - predict(model, newdata = data_tst))^ 2))
}

get_perc_err = function(model, k) {
  actual = data_tst[, k]
  predicted = predict(model, newdata = data_tst)
  100 * mean((abs(actual - predicted)) / actual)
}
```
